Query Optimization for Dynamic Imputation

### 2017-07-10 paper/experiments.tex
Edits

\textbf{Accuracy vs. Base-table Imputation (COUNT)}\quad
We also calculate the number of tuples used to produce each aggregate output.
The count fraction columns in \Cref{table:smape} show the number of tuples in the aggregate for $\alpha = 0$ and $\alpha = 1$ as a fraction of the number of tuples used when running the query on the imputed base table.
This shows that when optimizing for performance, not quality, many tuples are dropped due to
insertion of $\delta$ operators.
Even in cases where the SMAPE reduction from $\alpha = 1$ and $\alpha = 0$ is small
(\Cref{q2} and \Cref{q6}) the tuple count is significantly different.
In these cases, the aggregate value is not significantly impacted by the missing data. In
particular, if values are missing completely at random,
the aggregate should not be affected.
However, if the missing data is biased then the aggregate will have a significant error.
This highlights a challenge for a user handling data imputation traditionally: it is unclear if the missing data will have a large or small negative impact on their analysis until they have paid the cost of running it.

@@

\textbf{Alternate Imputation Methods}\quad
We also experiment with mean-value imputation and hot deck imputation
(\Cref{sec:imputation}). In our implementation of mean-value imputation, to facilitate
non-blocking operation within the iterator model, we estimate the column mean from base
table histograms. Therefore, imputed values are identical no matter where the operator is
placed in the query plan (relative errors are $0$). Runtimes are in the range of
\runtimemeanlow{} to \runtimemeanhigh{} for queries using \ProjectName{}, as compared to
\runtimemeanlowbase{} to \runtimemeanhighbase{} when imputing on the base table.  Here, the
time cost of the imputation is low no matter how many tuples need to be imputed.

>>>>

\textbf{Accuracy vs. Base-table Imputation (COUNT)}\quad
We calculate the number of tuples used to produce each aggregate output.
The count fraction columns in \Cref{table:smape} show the number of tuples in the aggregate for $\alpha = 0$ and $\alpha = 1$ as a fraction of the number of tuples used when running the query on the imputed base table.
This shows that when optimizing for performance, not quality, many tuples are dropped due to
insertion of $\delta$ operators.
Even in cases where the SMAPE reduction from $\alpha = 1$ and $\alpha = 0$ is small
(\Cref{q2} and \Cref{q6}) the tuple count is significantly different.
In these cases, the aggregate value is not significantly impacted by the missing data. In
particular, if values are missing completely at random,
the aggregate should not be affected.
However, if the missing data is biased then the aggregate will have a significant error.
This highlights a challenge for a user handling data imputation traditionally: it is unclear if the missing data will have a large or small negative impact on their analysis until they have paid the cost of running it.

@@

\textbf{Alternate Imputation Methods}\quad
We experiment with mean-value imputation and hot deck imputation
(\Cref{sec:imputation}). In our implementation of mean-value imputation, to facilitate
non-blocking operation within the iterator model, we estimate the column mean from base
table histograms. Therefore, imputed values are identical no matter where the operator is
placed in the query plan (relative errors are $0$). Runtimes are in the range of
\runtimemeanlow{} to \runtimemeanhigh{} for queries using \ProjectName{}, as compared to
\runtimemeanlowbase{} to \runtimemeanhighbase{} when imputing on the base table.  Here, the
time cost of the imputation is low no matter how many tuples need to be imputed.
